Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German

### 2024-12-02 paper/revision2/ResponseToReviewers.tex

\begin{quote}
    \Paste{thisdesign2}
\end{quote}

\begin{quote}
    \Paste{thisdesign3}
\end{quote}

\subsection*{Comment E.3}
\begin{quote}
``\textbf{Reviewer 2's additional point 1 that asks whether you have changed that particular analysis}.''
\end{quote}

\subsubsection*{Response to E.3}
Yes, as we had explained in the response to reviewers on August 29, 2024, all analyses were changed in the first round of revisions. For a fair comparison of our results to the results of the previous studies, especially Van Dyke (2007), it was necessary to compute Bayes factors using the Savage-Dickey ratio (in the original version of the manuscript we had used bridge sampling). Furthermore, for the Van Dyke (2007) data, the first on this design, it only made sense to use symmetrical priors which are agnostic regarding the sign of the effects. For consistency, we then used symmetrical priors for all our analyses (compared to the directional priors which we had used originally).

These changes did not change the results qualitatively, but led to some numerical changes which probably piqued the reviewer's interest.

@@

\subsubsection*{Response to R2.2}
The focus of the present study was to investigate retrieval interference due to syntactic and semantic similarity. The finding that the reading times were primarily affected by encoding interference was unexpected (but see \citeauthor{mertzen}'s (\citeyear{mertzen}) discussion). Given that the present manuscript contains two large-sample experiments and computational modeling (and its substantial length), we think that it is outside the scope of this manuscript to provide an in-depth discussion of a mechanism of encoding interference. Developing a complete treatment of encoding interference would require considerably more computational modeling, which would further increase the length and complexity of this paper. In our opinion, such a model should be developed, but that would need to be a stand-alone article, which would also require more experimental studies to validate and test competing models' predictions. For a recent attempt that we made to try to implement different versions of feature distortion, a phenomenon that is related to encoding interference, see:

Himanshu Yadav, Garrett Smith, Sebastian Reich, and Shravan Vasishth. Number feature distortion modulates cue-based retrieval in reading. Journal of Memory and Language, 129, 2023.

Another interesting model of encoding interference is

Smith, Garrett, Julie Franck, and Whitney Tabor. ``Encoding interference effects support self-organized sentence processing." Cognitive Psychology 124 (2021): 101356.

We have now cited this paper on pages \pageref{revised_SPR} and \pageref{smith2}:

\Paste{smith1}
\\

\Paste{smith2}\\

So, in summary, a proper treatment of encoding interference would require a modeling effort at the same scale as was carried out in \citet{Yadavetal2022}. Such a modeling effort, although very worthwhile, would take several years' of work and is therefore far beyond the scope of the present work.

### 2024-12-16 paper/revision2/elsarticle-template-harv.tex
Updated via the Open Science Framework

%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references

\documentclass[review,preprint,12pt,authoryear,floatsintext]{elsarticle}

@@

\begin{frontmatter}
\title{Do syntactic and semantic similarity lead to interference effects?\\
Evidence from self-paced reading and event-related potentials using German}
%% Title, authors and addresses

@@

\begin{abstract}
%% Text of abstract
\Copy{abstract}{
Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed. If these retrieval cues match with not only the \textcolor{blue}{subject} but also with those on other nouns in the sentence, then processing difficulty arises at the verb. This difficulty in identifying the correct dependent is called similarity-based interference. 
We present relatively large-sample self-paced reading and event-related potentials experiments using a well-established design to investigate interference due to syntactic and semantic cues in German. \textcolor{blue}{In this design, the syntactic cue \{+subject\} and the semantic cue \{+animate\} are manipulated.}
\textcolor{blue}{Bayes factors analyses showed} evidence for a semantic interference effect \textcolor{blue}{in both experiments}. 
Surprisingly, Bayes factors provided evidence against interference due to \textcolor{blue}{the syntactic cue \{+grammatical subject\} in this particular design in both experiments}. This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assumes that both syntactic and semantic cues play an equal role in retrieval. We show through computational modeling that cue-based retrieval \textcolor{blue}{will} also shows no syntactic interference \textcolor{blue}{in the present design} if the parser is assumed to keep track of which clause the subject occurs in. Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the cue-based retrieval model would exhibit patterns consistent with the observed patterns in our data. 
}
\end{abstract}

@@

The SPR study, hereafter Experiment 1a, was initially designed so that participants read \textcolor{blue}{half of the items (1-60) in the first session and the other half of the items (61-120) in the second session on another day.} The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall. The adaptation effect is theoretically interesting per se \citep{fine2013rapid}; but it was problematic for our design because the average effect would then be confounded by adaptation. For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=570) were either shown items 1-60 or items 61-120 (but not both). In the Results section, we present the pooled data (N=774) from E1a and E1b.
As mentioned above, because the ERP method requires a large number of items, all the ERP data come from participants who completed both sessions.

@@

In Experiment 1a, participants were re-invited for session 2 after they completed session 1. The experimental sessions were separated by 1 - 20 days. The procedure of the sessions was identical. In session 1, the experimental items 1-60 were presented interspersed with 40 filler sentences. In session 2, the experimental items 61-120 were presented with another set of 40 filler sentences. {\Copy{fillers}{\textcolor{blue}{The fillers were included to mask the critical manipulation. They} were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety\textcolor{blue}{, e.g., by including modifiers and different types of subordinate clauses} (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich, während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers} Each session lasted approximately 25 minutes.

@@

\Copy{trim}{{\textcolor{blue}{Self-paced reading times below 150 ms and above 3000 ms were excluded before analysis because these overly short or long reaction times were likely caused by inattentively performing the task. The procedure excluded 4.9\,\% of the data.}}}\label{trim} Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}. The critical word for analyses was the verb \textit{trank} `drank' in (\ref{ex:materials}), constituting the retrieval site. Because previous work found effects in the pre-critical and post-critical region, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed. 
The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction. In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task. The trial id of the 100 trials per session was rescaled to span from 0 to 1. This was done to bring all fixed effects to a comparable scale which decreases the run time of the models. The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items.

@@

Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panel A)} and low \textcolor{blue}{(panel B)} syntactic interference because these conditions had partially different sentence structures.

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. \textcolor{blue}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled.jpg}
\end{sidewaysfigure}
\clearpage

\label{revised_SPR}\Copy{revised_SPR}{
It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. \Copy{smith1}{\textcolor{blue}{To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006, smith2021encoding}}}\label{smith1}. \textcolor{blue}{Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session. As mentioned earlier, this attenuation in reading times was likely due to adaptation to the SPR task in the second session.

\textcolor{blue}{Crucially, the reading times difference starting at the distractor} persisted in the following regions. The distractor and the immediately following regions differed between conditions, \textcolor{blue}{therefore we did not analyze their reading times. In our materials, there were two words between the distractor and the critical verb which were identical across conditions: the pre-pre-critical and pre-critical adverbs. The pre-pre-critical region was introduced in the present study to absorb potential effects caused by processing the clause boundary of the embedded clause, rendering it a problematic region for analyses. The pre-critical region was not adjacent to clause boundaries. The reading times of the pre-critical region were analyzed to investigate the reading times difference starting at the distractor. This difference starting at the distractor and persisting almost until the end of the sentence was problematic for the analysis and interpretation of reading times in the critical (and spill-over) region. Any effects that might be present there cannot be attributed clearly to the processing of the respective region. However, for comparability with previous work, we briefly also report analyses of the reading times of the critical and spill-over region.}

Figure \ref{fig:spr_posteriors} shows that \textcolor{blue}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions. The estimates of semantic interference were positive in all regions and the estimates of the interaction were mostly positive in all regions. The estimates of syntactic interference were less consistent across regions. They were almost centered around zero in the critical region, showed mostly positive values in the pre-critical region and fully positive values in the spill-over region.}
}

@@

\paragraph{Hypothesis testing using Bayes factors}
\textcolor{blue}{To complete the analysis of the effects in the pre-critical, critical and spill-over regions -- although they were confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}). There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5, Normal(0, 0.05): BF$_{10}$ = 1.3, Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).

@@

\subsection*{Discussion}
\Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{blue}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{However, reading times started to differ before this critical dependency could be formed. Starting at the distractor, which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors). Because the distractor differed between conditions, we did not analyze distractor reading times. Instead, to investigate the difference starting after the distractor was read, we investigate the pre-critical region, which was identical across conditions and preceded the critical verb. The pre-critical region serves as a proxy for any encoding interference effects that arose after the distractor was read. Bayes factors provided extremely strong evidence for a semantic interference effect in the pre-critical region. In contrast, there was evidence against syntactic interference and an interaction in that region. We refrain from interpreting effects in the critical and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes, i.e., memory retrieval. The reading times differences at the distractor and the pre-critical region are discussed below.}}

\subsubsection*{\textcolor{blue}{Reading times differences prior to retrieval}}\label{prior_to_retrieval}

\textcolor{blue}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence}). One might argue that this could be due to differences in word length. The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9). It seems unlikely that this small difference of one letter in word length caused such large and long-lasting effects, but of course one cannot rule out this possibility with complete certainty.

Another possible explanation for the reading time differences starting at the distractor could be \textcolor{blue}{that the distractors induced} a difference in plausibility between the high and low semantic interference conditions. To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \citep{pcibex}. Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}, repeated here as \ref{ex:hisyn_repeat} for convenience).

@@

To investigate whether this plausibility difference was driving the \textcolor{blue}{reading time difference between the semantic interference conditions}, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region. This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings. We used the same priors as presented in Table \ref{tab:spr_priors}. The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items. Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes. Although plausibility influenced reading times (higher plausibility led to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms). The Bayes factors provided evidence for semantic interference in this analysis under all priors except the most diffuse one (Normal(0, 0.01): BF$_{10}$ = 4.2,
Normal(0, 0.05): BF$_{10}$ = 2.4,
Normal(0, 0.1): BF$_{10}$ = 1.3). The interaction of semantic interference and plausibility was centered around zero. In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference \textcolor{blue}{starting at the distractor and lasting into later regions} was partially caused by reduced plausibility of the high semantic interference conditions. However, the plausibility difference was not the only driver of the reading times slowdown, i.e., \textcolor{blue}{high} semantic interference led to slower reading times independent of plausibility.

@@

\textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference.} \Copy{smith2}{The most likely explanation is encoding interference \cite[e.g.,][but see the potential confounds discussed above]{Oberauer_Kliegl_2006,smith2021encoding}.}\label{smith2} The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.

@@

\cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region. {\Copy{confound}{In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions \textcolor{blue}{(two vs.\ one embedded clause between subject and verb in the high vs.\ low syntactic interference conditions)}, and predictive processing effects.}}\label{confound} Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.

@@

Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region. However, in our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence}). The predictive-processing explanation would incorrectly predict that the effect begins at the first adverb, which is the first word of the verbal phrase. Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data. Crucially and as mentioned above, \textcolor{blue}{due to the long-lasting reading time difference starting at the distractor} it is not possible to pinpoint \textcolor{blue}{the driver of effects observed in later regions. The effects in the critical region (and spill-over region) might be} due to a retrieval initiated at the respective region, or still \textcolor{blue}{be} due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}.

\textcolor{blue}{In summary, the main finding of Experiment 1 is that the reading times were primarily affected by semantic encoding interference.} We turn next to the event-related potentials experiment.

@@

\label{modeling}{The \citet{Lewis2005} model is generally used to predict reading times. Recently, \cite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb. We found that the reading times at the critical verb were confounded by \textcolor{blue}{reading times differences starting earlier in the sentence}, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}. Because the Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions. However, the ERP data can be compared to the model's predictions; all that has to be changed is that the measurement is now on the microvolt scale instead of the millisecond scale. In order to model the ERP data, we therefore adjust the scaling factor $F$, which rescales the activation of items in memory to the relevant dependent measure. The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.

@@

\subsection*{\textcolor{blue}{The use of syntactic and semantic information}}

>>>>

\documentclass[12pt,authoryear,floatsintext]{elsarticle}

@@

\begin{frontmatter}
\title{Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German}

\author[add1]{Pia Schoknecht\corref{cor1}}
\ead{pia.schoknecht@uni-potsdam.de}
\author[add2]{Himanshu Yadav}
\author[add1]{Shravan Vasishth}

\cortext[cor1]{Corresponding author}

\address[add1]{University of Potsdam, Department of Linguistics, Karl-Liebknecht-Straße 24-25, 14476 Potsdam, Germany}
\address[add2]{Indian Institute of Technology Kanpur, Department of Cognitive Science, Kanpur 208016, India}

%% Title, authors and addresses

@@

\begin{abstract}
%% Text of abstract
\Copy{abstract}{
Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed. If these retrieval cues match with not only the \textcolor{black}{subject} but also with those on other nouns in the sentence, then processing difficulty arises at the verb. This difficulty in identifying the correct dependent is called similarity-based interference. 
We present relatively large-sample self-paced reading and event-related potentials experiments using a well-established design to investigate interference due to syntactic and semantic cues in German. \textcolor{black}{In this design, the syntactic cue \{+subject\} and the semantic cue \{+animate\} are manipulated.}
\textcolor{black}{Bayes factors analyses showed} evidence for a semantic interference effect \textcolor{black}{in both experiments}. 
Surprisingly, Bayes factors provided evidence against interference due to \textcolor{black}{the syntactic cue \{+grammatical subject\} in this particular design in both experiments}. This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assumes that both syntactic and semantic cues play an equal role in retrieval. We show through computational modeling that cue-based retrieval \textcolor{black}{will} also show no syntactic interference \textcolor{black}{in the present design} if the parser is assumed to keep track of which clause the subject occurs in. Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the cue-based retrieval model would exhibit patterns consistent with the observed patterns in our data. 
}
\end{abstract}

@@

The SPR study, hereafter Experiment 1a, was initially designed so that participants read \textcolor{black}{half of the items (1-60) in the first session and the other half of the items (61-120) in the second session on another day.} The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall. The adaptation effect is theoretically interesting per se \citep{fine2013rapid}; but it was problematic for our design because the average effect would then be confounded by adaptation. For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=570) were either shown items 1-60 or items 61-120 (but not both). In the Results section, we present the pooled data (N=774) from E1a and E1b.
As mentioned above, because the ERP method requires a large number of items, all the ERP data come from participants who completed both sessions.

@@

In Experiment 1a, participants were re-invited for session 2 after they completed session 1. The experimental sessions were separated by 1 - 20 days. The procedure of the sessions was identical. In session 1, the experimental items 1-60 were presented interspersed with 40 filler sentences. In session 2, the experimental items 61-120 were presented with another set of 40 filler sentences. {\Copy{fillers}{\textcolor{black}{The fillers were included to mask the critical manipulation. They} were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety\textcolor{black}{, e.g., by including modifiers and different types of subordinate clauses} (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich, während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers} Each session lasted approximately 25 minutes.

@@

\Copy{trim}{{\textcolor{black}{Self-paced reading times below 150 ms and above 3000 ms were excluded before analysis because these overly short or long reaction times were likely caused by inattentively performing the task. The procedure excluded 4.9\,\% of the data.}}}\label{trim} Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}. The critical word for analyses was the verb \textit{trank} `drank' in (\ref{ex:materials}), constituting the retrieval site. Because previous work found effects in the pre-critical and post-critical region, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed. 
The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction. In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task. The trial id of the 100 trials per session was rescaled to span from 0 to 1. This was done to bring all fixed effects to a comparable scale which decreases the run time of the models. The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items.

@@

Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{black}{(panel A)} and low \textcolor{black}{(panel B)} syntactic interference because these conditions had partially different sentence structures.

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. \textcolor{black}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled.jpg}
\end{sidewaysfigure}
\clearpage

\label{revised_SPR}\Copy{revised_SPR}{
It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. \Copy{smith1}{\textcolor{black}{To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006, smith2021encoding}}}\label{smith1}. \textcolor{black}{Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session. As mentioned earlier, this attenuation in reading times was likely due to adaptation to the SPR task in the second session.

\textcolor{black}{Crucially, the reading times difference starting at the distractor} persisted in the following regions. The distractor and the immediately following regions differed between conditions, \textcolor{black}{therefore we did not analyze their reading times. In our materials, there were two words between the distractor and the critical verb which were identical across conditions: the pre-pre-critical and pre-critical adverbs. The pre-pre-critical region was introduced in the present study to absorb potential effects caused by processing the clause boundary of the embedded clause, rendering it a problematic region for analyses. The pre-critical region was not adjacent to clause boundaries. The reading times of the pre-critical region were analyzed to investigate the reading times difference starting at the distractor. This difference starting at the distractor and persisting almost until the end of the sentence was problematic for the analysis and interpretation of reading times in the critical (and spill-over) region. Any effects that might be present there cannot be attributed clearly to the processing of the respective region. However, for comparability with previous work, we briefly also report analyses of the reading times of the critical and spill-over region.}

Figure \ref{fig:spr_posteriors} shows that \textcolor{black}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions. The estimates of semantic interference were positive in all regions and the estimates of the interaction were mostly positive in all regions. The estimates of syntactic interference were less consistent across regions. They were almost centered around zero in the critical region, showed mostly positive values in the pre-critical region and fully positive values in the spill-over region.}
}

@@

\paragraph{Hypothesis testing using Bayes factors}
\textcolor{black}{To complete the analysis of the effects in the pre-critical, critical and spill-over regions -- although they were confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}). There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak \textcolor{black}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5, Normal(0, 0.05): BF$_{10}$ = 1.3, Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).

@@

\subsection*{Discussion}
\Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{black}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{black}{However, reading times started to differ before this critical dependency could be formed. Starting at the distractor, which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors). Because the distractor differed between conditions, we did not analyze distractor reading times. Instead, to investigate the difference starting after the distractor was read, we investigate the pre-critical region, which was identical across conditions and preceded the critical verb. The pre-critical region serves as a proxy for any encoding interference effects that arose after the distractor was read. Bayes factors provided extremely strong evidence for a semantic interference effect in the pre-critical region. In contrast, there was evidence against syntactic interference and an interaction in that region. We refrain from interpreting effects in the critical and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes, i.e., memory retrieval. The reading times differences at the distractor and the pre-critical region are discussed below.}}

\subsubsection*{\textcolor{black}{Reading times differences prior to retrieval}}\label{prior_to_retrieval}

\textcolor{black}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence}). One might argue that this could be due to differences in word length. The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9). It seems unlikely that this small difference of one letter in word length caused such large and long-lasting effects, but of course one cannot rule out this possibility with complete certainty.

Another possible explanation for the reading time differences starting at the distractor could be \textcolor{black}{that the distractors induced} a difference in plausibility between the high and low semantic interference conditions. To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \citep{pcibex}. Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}, repeated here as \ref{ex:hisyn_repeat} for convenience).

@@

To investigate whether this plausibility difference was driving the \textcolor{black}{reading time difference between the semantic interference conditions}, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region. This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings. We used the same priors as presented in Table \ref{tab:spr_priors}. The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items. Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes. Although plausibility influenced reading times (higher plausibility led to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms). The Bayes factors provided evidence for semantic interference in this analysis under all priors except the most diffuse one (Normal(0, 0.01): BF$_{10}$ = 4.2,
Normal(0, 0.05): BF$_{10}$ = 2.4,
Normal(0, 0.1): BF$_{10}$ = 1.3). The interaction of semantic interference and plausibility was centered around zero. In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference \textcolor{black}{starting at the distractor and lasting into later regions} was partially caused by reduced plausibility of the high semantic interference conditions. However, the plausibility difference was not the only driver of the reading times slowdown, i.e., \textcolor{black}{high} semantic interference led to slower reading times independent of plausibility.

@@

\textcolor{black}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference.} \Copy{smith2}{The most likely explanation is encoding interference \cite[e.g.,][but see the potential confounds discussed above]{Oberauer_Kliegl_2006,smith2021encoding}.}\label{smith2} The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.

@@

\cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region. {\Copy{confound}{In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions \textcolor{black}{(two vs.\ one embedded clause between subject and verb in the high vs.\ low syntactic interference conditions)}, and predictive processing effects.}}\label{confound} Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.

@@

Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region. However, in our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence}). The predictive-processing explanation would incorrectly predict that the effect begins at the first adverb, which is the first word of the verbal phrase. Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data. Crucially and as mentioned above, \textcolor{black}{due to the long-lasting reading time difference starting at the distractor} it is not possible to pinpoint \textcolor{black}{the driver of effects observed in later regions. The effects in the critical region (and spill-over region) might be} due to a retrieval initiated at the respective region, or still \textcolor{black}{be} due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}.

\textcolor{black}{In summary, the main finding of Experiment 1 is that the reading times were primarily affected by semantic encoding interference.} We turn next to the event-related potentials experiment.

@@

\label{modeling}{The \citet{Lewis2005} model is generally used to predict reading times. Recently, \cite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb. We found that the reading times at the critical verb were confounded by \textcolor{black}{reading times differences starting earlier in the sentence}, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}. Because the Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions. However, the ERP data can be compared to the model's predictions; all that has to be changed is that the measurement is now on the microvolt scale instead of the millisecond scale. In order to model the ERP data, we therefore adjust the scaling factor $F$, which rescales the activation of items in memory to the relevant dependent measure. The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.

@@

\subsection*{\textcolor{black}{The use of syntactic and semantic information}}
